Predicting the effects of character displacement on food-web dynamics

### 2019-03-07 manuscript_ECD_model.Rmd
edit title page and abstract

\vspace*{0.1cm} 
\begin{center} \LARGE Food-web dynamics under ecological character displacement \end{center}

@@

\noindent 1. University of Zurich, Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich, Switzerland;

@@

*Keywords*: competition; eco-evolutionary dynamics; consumer-resource interactions; adaptive radiation; ecological community

@@

*Manuscript type*: Note.

@@

Ecological character displacement is an adaptive process, often enhancing coexistence between consumers by reducing competition for shared resources. 
This focus on competition has undermined the fact that consumer-resource interactions ultimately mediate the process of character displacement. 
As a consequence, the effects of character displacement on consumer-resource interactions and food-web dynamics have not been critically examined.
Here, I study a model of two consumers competing for two shared resources to examine how character displacement in consumers influences resource densities and the resilience of food webs to perturbations.
To test the generality of these effects, I examined them under different assumptions of evolutionary tradeoffs, consumer foraging behavior, and resource distributions.
My general finding was that, whenever character displacement increased the strength of consumer-resource interactions, it resulted in lower resource densities and less resilient food webs. 
Character displacement always strengthened consumer-resource interactions whenever consumers were competing for resources that occured in distinct habitats. 
This occurred under different evolutionary tradeoffs and in both simple and more realistic foraging scenarios.
While character displacement did not always increase the strength of consumer-resource interactions, these contingent effects only occurred under the simplest and least realistic foraging scenario. 
Taken together, my results indicate that the adaptive process of character displacement often comes at the ecological cost of making food webs less resilient to perturbations. 
They also suggest that food-web and competition theory can give complimentary insights to the ecological consequences of adaptation.

>>>>

\vspace*{0.1cm} 
\begin{center} \LARGE Ecological character displacement destabilizes food webs \end{center}
<!--\begin{center} \LARGE Food-web dynamics under ecological character displacement \end{center}-->

@@

\noindent 1. University of Zurich, Department of Evolutionary Biology and Environmental Studies, 8057 Zurich, Switzerland;

@@

*Keywords*: competition; eco-evolutionary dynamics; consumer-resource interactions; adaptation; ecological community; stability

@@

*Manuscript type*: Note. <!-- consider e-Article, depending on length -->

@@

Ecological character displacement is an adaptive process, often enhancing coexistence between consumers by reducing competition for shared resources. 
This competition perspective undermine the fact that consumer-resource interactions ultimately mediate the process of character displacement. 
As a consequence, the effects of character displacement on consumer-resource interactions and food-web dynamics have been ignored.
Here, I study a model of two consumers competing for two shared resources to examine how character displacement in consumers affects resource densities and the resilience of food webs to perturbations.
I studied these effects under different assumptions of evolutionary tradeoffs, consumer foraging behavior, and resource distributions.
I found that character displacement always strengthened consumer-resource interactions whenever consumers competed for resources that occured in distinct habitats.
This increase in interaction strength resulted in lower resource abundances and less resilient food webs.
This occurred under different evolutionary tradeoffs and in both simple and more realistic foraging scenarios.
While character displacement did not always increase the strength of consumer-resource interactions, these contingent effects only occurred under the simplest and least realistic foraging scenario. 
Taken together, my results show that the adaptive process of character displacement often comes at the ecological cost of decreasing food-web resilience. 
They also highlight that food-web and competition theory can give complimentary insights to the ecological consequences of adaptation.
